---
title: "gt and reactable examples"
output:
  html_notebook: default
---

# The `gt` package for static tables

The `gt` package is a brilliant new R package for building static tables. `gt` gives you a ton of flexibility in how you build your table, and works well with the `tidyverse`. You can learn more about `gt` at [gt.rstudio.com](https://gt.rstudio.com/).


```{r setup, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = NA)

library(gt)
library(reactable)
library(tidyverse)
library(tidyquant)
```


Let's look at an example using some stock data. We'll use the `tidyquant` package to load some stock data and calculate returns. For this example, we'll use the FAANG stocks.

```{r}
symbols <- c("FB", "AAPL", "AMZN", "NFLX", "GOOG")

prices <- symbols %>% 
  tq_get(get = "stock.prices", from = "2011-01-01") %>% 
  group_by(symbol) %>% 
  arrange(date)

prices
```

Next, let's calculate our daily returns for each stock. We'll filter by date to just last week, and use this table for printing.

```{r}
stock_table <- prices %>% 
  tq_mutate(adjusted, periodReturn, period = "daily", col_rename = "returns") %>% 
  filter_by_time(date, "2021-01-11", "2021-01-15") %>% 
  select(symbol, date, adjusted, returns, volume) %>%
  arrange(date, symbol)

stock_table
```

Finally, let's assemble our `gt` table and set a few options.

```{r}
stock_table %>% 
  gt() %>% 
  tab_header(
    title = md("**FAANG** Stock Performance"),
    subtitle = glue::glue("{min(stock_table$date)} to {max(stock_table$date)}")
  ) %>% 
  fmt_date(
    columns = vars(date),
    date_style = 3
  ) %>% 
  fmt_currency(
    columns = vars(adjusted)
  ) %>% 
  fmt_percent(
    columns = vars(returns)
  ) %>% 
  fmt_number(
    columns = vars(volume),
    suffixing = TRUE
  ) %>% 
  tab_options(
   row_group.background.color  = "#fffcad"
  ) %>% 
  data_color(
    columns = vars(returns),
    colors = scales::col_numeric(
      palette = c("#ff1100", "#ffffff", "#00ff00"),
      domain = c(-.05, .05)
    )
  ) %>% 
  tab_source_note(
    source_note = md("Data from [Yahoo Finance](https://finance.yahoo.com/) via the [tidyquant](https://business-science.github.io/tidyquant/) R package.")
  ) %>% 
  tab_options(
    table.width = pct(90)
  )
```


# The `reactable` package for interactive tables

The [`reactable`](https://glin.github.io/reactable/index.html) package leverages the React and React Table libraries to allow you to build extremely customizable, highly interactive tables.

Reactable provides a ton of nice interactivity "for free", but really shines if you know some CSS and JavaScript.

Let's look at our same stock table, but make it interactive with reactable.

```{r}
options(reactable.theme = reactableTheme( # here is a dark theme
  color = "hsl(233, 9%, 87%)",
  backgroundColor = "hsl(233, 9%, 19%)",
  borderColor = "hsl(233, 9%, 22%)",
  stripedColor = "hsl(233, 12%, 22%)",
  highlightColor = "hsl(233, 12%, 24%)",
  inputStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  selectStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonHoverStyle = list(backgroundColor = "hsl(233, 9%, 25%)"),
  pageButtonActiveStyle = list(backgroundColor = "hsl(233, 9%, 28%)")
))

# options(reactable.theme = NULL) # run this to go back to default theme

stock_table %>%
  reactable(groupBy = "symbol",
            sortable = TRUE,     # sort by columns? defaults to true
            resizable = TRUE,    # resizable columns?
            borderless = FALSE,  # remove inner borders?
            outlined = TRUE,     # add an outer border?
            striped = TRUE,      # zebra striping for rows
            filterable = TRUE,   # enable column filtering
            searchable = TRUE,   # enable global table searching?
            highlight = TRUE,    # highlight table rows on hover?
            pagination = TRUE,  # whether to paginate rows
            columns = list(
                    adjusted = colDef(format = colFormat(currency = "USD")),
                    returns = colDef(format = colFormat(percent = TRUE, digits = 2)),
                    volume = colDef(aggregate = "sum", format = colFormat(digits = 0))),
            details = function(index){
              htmltools::div(
                "The adjusted closing price for ", stock_table[index, ]$symbol, " on ", stock_table[index, ]$date, " was $", round(stock_table[index, ]$adjusted, 2), ", representing a ", round(stock_table[index, ]$returns*100, 2), "% change from the previous value."
              )
            },
            defaultColDef = colDef(footer = function(values) {
              if (!is.numeric(values)) return()
              sparkline(values)
              })
  )
```

To learn more, check out some of the demos at https://glin.github.io/reactable/index.html.
